% The Ouroboros Thesis v22 -- Convergence
% Reconstructed LaTeX source for arXiv submission (from PDF + markdown).
% Compile: pdflatex thesis_v22 ; bibtex thesis_v22 ; pdflatex x2
% License: CC BY 4.0.  Code: Apache-2.0.
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{xcolor}

% --- Greek/math macros to avoid Type-3 fonts and keep ASCII source ---
\newcommand{\Lam}{\ensuremath{\Lambda}}
\hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, citecolor=blue}

\title{The Ouroboros Thesis v22 --- Convergence\\[2pt]
\large An Honest, Audit-Ready Convergence of the \Lam-Aggregator Uniqueness Chain,
Mechanism Truthfulness, and Sim-to-Real Doctrine Transfer}
\author{Stephen P. Lutar Jr.\\ SZL Holdings, Inc.\\
\texttt{ORCID 0009-0001-0110-4173}}
\date{2026-06-03}

% MSC2020: 39B22 (primary), 26E60, 91B26, 68V20, 94A17.  ACM: F.4.1, F.2.0.

\begin{document}
\maketitle

\begin{center}\small\fbox{\parbox{0.95\linewidth}{\centering
Doctrine v11 LOCKED --- 749 declarations $\cdot$ 14 unique axioms $\cdot$ 163 sorries @ \texttt{c7c0ba17}
$\cdot$ A5 merged (v11.1 in flight: 794/14/191, axioms unchanged) $\cdot$
\Lam-aggregator uniqueness is \textbf{Conjecture 1 --- NOT a theorem} $\cdot$ SLSA L1 honest · L2 build-attested (Rekor) · L3+ roadmap (NOT L3).}}
\end{center}

\begin{abstract}
v22 (``Convergence'') consolidates the May--June 2026 formal-verification advances into the canonical
thesis line. Its central correction: the claim that axioms A1--A4 force the weighted geometric mean is
\emph{false}. The asymmetric mean $\Phi(x_1,x_2)=x_1^{2/3}\,x_2^{1/3}$ satisfies A1--A4 yet differs
from \Lam{} and fails permutation invariance. We add A5 (permutation invariance) as a \emph{structure
field} on \texttt{LutarAxioms} --- not a new axiom --- keeping the unique-axiom count at 14, and report
the partial closure of the $n$-dimensional Cauchy functional-equation chain that, when complete, would
discharge \Lam-uniqueness. It is \emph{not} complete on \texttt{main}; \Lam{} therefore remains
\textbf{Conjecture 1}. We additionally report VCG mechanism truthfulness (proven on branch), the
achievement of SLSA L2 build provenance (verified on all five GHCR images), the Round 10--11 frontier
formalizations, and a Sim-to-Real doctrine-transfer benchmark modeled on the Walrus physical foundation
model that measures a mean doctrine $\alpha$-gap of $0.10$ across five unseen compliance regimes. We
claim only what is mechanically checked or empirically measured.
\end{abstract}

\section{Recent advances (the substance of v22)}

\subsection{A5 axiom merge --- the A1--A4 uniqueness gap, corrected (MERGED, PR \#148)}
The historical ``Theorem~1 --- \Lam{} is the unique aggregator under A1--A4'' was \emph{incorrect}. A
review of 13 published results on quasi-arithmetic and symmetric means confirms A1--A4 are insufficient
to force the geometric mean~\cite{kolmogorov1930,nagumo1930,aczel1966,hardy1934,voorneveld2008}.

\paragraph{Counterexample (verified).} $\Phi(x_1,x_2)=x_1^{2/3}\,x_2^{1/3}$ satisfies homogeneity (A2),
boundedness (A4) and the A1/A3 conditions, yet $\Phi(2,1)=2^{2/3}\neq 2^{1/3}=\Phi(1,2)$, so
permutation invariance fails and $\Phi\neq\Lambda_2$.

\paragraph{Fix (landed on \texttt{main}).} Add an \texttt{IsPermutationInvariant} predicate and an
\textbf{A5 structure field} to \texttt{LutarAxioms}. \texttt{Lambda\_A5\_perm\_invariant} is sorry-free
(\texttt{Equiv.prod\_comp}/\texttt{Fintype.prod\_equiv}). Because A5 is a structure field, the
unique-axiom count stays \textbf{14}; the live corpus moves to 794 declarations / 14 axioms / 191 sorries.

\subsection{The Cauchy\_ND uniqueness chain --- partial closure (IN REVIEW)}
With A5 in place, \Lam-uniqueness reduces to an $n$-dimensional Cauchy functional-equation chain.

\begin{center}\small
\begin{tabular}{@{}llll@{}}
\toprule
Branch & PR & Status & Note \\
\midrule
Topology (monotone $\to$ continuity bridge) & \#175 & in review & Landed \emph{true forms}; refused to fake-prove \\
Functional analysis (\texttt{multiplicative\_monotone\_isPow}) & \#173 & in review & 1 honest sorry on the $t=0$ degenerate case \\
Symmetric (exponents $\alpha_i=1/k$) & \#174 & in review & Closed with A5 dependency \\
\bottomrule
\end{tabular}
\end{center}

\noindent The combined chain \texttt{A5 + Cauchy + topology + symmetric} is the full \Lam-uniqueness
chain. \textbf{It is not yet complete on \texttt{main}} (three PRs open, one residual honest sorry).
\textbf{Therefore \Lam{} stays Conjecture~1.} We will elevate \Lam{} to Theorem~1 \emph{only} when every
Cauchy\_ND sorry closes on \texttt{main} and Lake CI is green.

The proof route follows Acz\'el (1966, Thm~5.1)~\cite{aczel1966} and Cauchy (1821)~\cite{cauchy1821}:
the axis slice $f_i(t)=\Lambda(1,\dots,t,\dots,1)$ is monotone (A1) with $f_i(1)=1$ (A3); the log-space
map $g_i(u)=\log f_i(e^u)$ is additive and monotone, hence linear, giving $f_i(t)=t^{\alpha_i}$; A5 forces
$\alpha_i=\alpha_j$; A3 pins $\alpha=1/k$; reassembly yields $\Lambda(x)=\bigl(\prod_i x_i\bigr)^{1/k}$.

\subsection{VCG mechanism truthfulness (IN REVIEW, PR \#172)}
Both VCG obligations are closed on branch: \texttt{vcgDominantStrategyTruth} (truthful bidding is a
dominant strategy) and \texttt{vcgIndividualRationality} (participation never yields negative utility),
via Mathlib's \texttt{Finset.exists\_max\_image} and \texttt{add\_sum\_erase}.

\subsection{SLSA L1 honest · L2 build-attested (Rekor) · L3+ roadmap (ATTESTED)}
Build-provenance posture upgraded to L1 + L2; all 5/5 flagship GHCR images (a11oy, sentra, amaru,
killinchu, rosie) empirically verified with \texttt{slsa-verifier}. SLSA L3 is \emph{not} claimed ---
hardened/isolated builders remain roadmap.

\subsection{Innovation Rounds 10--11 (IN REVIEW / IN FLIGHT)}
Round 10 instilled frontier formalizations (all in review): Physics (\#177; Noether, Liouville,
Hamiltonian, entropy, A5-from-gauge), Quantum (\#176; post-quantum signatures, Holevo
bound~\cite{holevo1973}, Kitaev, ZK, no-cloning$\to$A5), CS (\#178; Byzantine quorum $n\ge 3f+1$,
FLP impossibility~\cite{flp1985}, CAP), Crypto (\#179; DSSE EUF-CMA, Rekor Merkle
inclusion~\cite{rfc6962}, Fulcio chain, BLS aggregation~\cite{bls2001}), distributed systems (PR
pending), Round~9 anatomy (\#170). Round~11 (formula frontier) is in flight.

\subsection{Sim-to-Real doctrine transfer --- Walrus parallel (DESIGN + PARTIAL EMPIRICAL)}
Modeled on the Walrus physical foundation model~\cite{walrus2025}, we treat the locked doctrine kernel
as a pretraining prior and a customer's few-shot receipt set as fine-tuning data, and define the doctrine
$\alpha$-gap $=\lvert\text{OOD acc.}-\text{in-distribution acc.}\rvert$. On a live $N{=}60$ run against
sentra and a11oy:

\begin{center}\small
\begin{tabular}{@{}lcc@{}}
\toprule
Regime & Accuracy & $\alpha$-gap \\
\midrule
R0 control & 1.00 & --- \\
R1 adversarial & 0.50 & 0.50 \\
R2 cross-jurisdictional & 1.00 & 0.00 \\
R3 multimodal & 1.00 & 0.00 \\
R4 temporal-drift & 1.00 & 0.00 \\
R5 low-data & 1.00 & 0.00 \\
\midrule
\textbf{Mean} & --- & \textbf{0.10} \\
\bottomrule
\end{tabular}
\end{center}

\noindent Four of five unseen regimes transfer perfectly; adversarial transfers only partially (signature
blocklist misses semantically novel attacks). We claim the architecture \emph{admits} sim-to-real
transfer --- not that it matches physical-FM accuracy. This is a single $N{=}60$ run reported as
partial empirical evidence, not a generalization guarantee.

\section{Doctrine attestation (verbatim)}
\begin{itemize}\setlength\itemsep{0pt}
  \item Doctrine version: v11 LOCKED (v11.1 in flight, post-A5).
  \item Declarations 749 (pinned @ \texttt{c7c0ba17}); 794 post-A5 live.
  \item Unique axioms: \textbf{14} (unchanged by A5 --- structure field).
  \item Sorries: 163 pinned (112 baseline + 51 Putnam); 191 post-A5 live.
  \item \Lam{} status: \textbf{Conjecture 1 --- NOT a theorem.}
  \item Supply chain: SLSA L1 honest · L2 build-attested (Rekor) · L3+ roadmap; NOT L3.
  \item Section 889 vendors: Huawei, ZTE, Hytera, Hikvision, Dahua (exactly 5).
\end{itemize}

\section{Honest posture}
\begin{itemize}\setlength\itemsep{0pt}
  \item \Lam-aggregator uniqueness is Conjecture~1, conditional on the Cauchy\_ND chain closing on \texttt{main}.
  \item A5 is a structure field, not a new axiom; axiom count remains 14.
  \item Reed--Solomon $\neq$ holographic. Event-sourcing $\neq$ time travel. Physics analogies are scaffolding.
  \item Quechua organ names are brand naming, not prior-art or cultural claims. No mystical claims.
  \item In-review PRs (\#170, \#172--\#179) are labeled as such and not presented as landed theorems.
\end{itemize}

\paragraph{Citation.} S.~P.~Lutar~Jr., ``The Ouroboros Thesis v22 --- Convergence,'' SZL Holdings,
Concept DOI \href{https://doi.org/10.5281/zenodo.19944926}{10.5281/zenodo.19944926}, 2026. The v22
version DOI is minted by Zenodo on the \texttt{paper-v22-1.0.0} release.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
